SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher

### 2018-11-19 back.tex
Updated Sections: Intro, Back, Method, Results

%% NOTE: The below paper is literally the same idea as ours. But poor implementation and no good results. How much to bash this paper?
Finally, the paper on Data Cache Prefetching with Perceptron
Learning~\cite{BadPerc} talks about the idea of two step prefetching.
The first step is an existing baseline prefetcher while the second
step is a perceptron based throttler.  The paper has lots of
limitations in terms of design and implementation.  That reflects in
the fact that it did not lead to any significant performance gain over
even the basic prefetchers like the Stride Prefetcher~\cite{Stride} and
the Markov Prefetcher~\cite{Markov}.

\subsection{Perceptron Learning in Architecture}
\label{Background-Perceptron}
Perceptron learning for computer architecture design has been around
for a while.  It was first popularized by Jimenez \textit{et. al.} for
doing branch prediction~\cite{Perc_Branch}. In it's original
implementation, the Program Counter of the instruction looking for a
branch prediction would index into a given table of perceptron
weights.  The retrieved weights would then be multiplied with the
history of branch outcomes stored in the History Register (which is a
feature) in a dot product fashion.  The sum obtained is thresholded at
0 \textit{i.e.}, if the prediction y\textsubscript{out} >= 0, the
prediction is made as \textit{true}, or branch taken.  Else if the sum
is < 0, the prediction is made as \textit{false}.

The structure that we use is derived from the model introduced in
Piecewise Linear Branch Prediction~\cite{Piece_Linear}.  Here the
feature itself is used to index into a hash of perceptron weights.
This form of indexing makes sure that the hyper-plane learned by the
perceptron weights is able to differentiate between linearly
inseparable outcomes.  \textit{<Is it true? Not sure>}

Perceptron Learning for Reuse Prediction~\cite{Perc_Reuse} extended the
hash perceptron architecture to perform prediction in context of cache
replacement.  Perceptron layer can learn to correlate the cache
replacement behaviour with a variety of features.  Each feature has
its dedicated table of perceptron weights which it indexes into.
Hence number of tables equals the number of features.  Once the
perceptron weights from all the tables are obtained, they're summed
and thresholded based on a pre-decided value.  This led to a
prediction mechanism that is highly accurate, has quick learning rate
and is adaptable to changing program behaviour.  The features
introduced in that paper were derived from Program Counters of the
current and the last few instructions and from the memory address of the
current access.  Even with such basic features, high level of
correlation could be established by perceptrons.

@@

\subsection{Signature Path Prefetcher Architecture}
\label{Background-SPP}
Since the work we have done uses SPP~\cite{SPP} as the underlying prefetcher, we are
presenting a brief recap of the ideas introduced in the original
paper, especially with an eye towards their relevance in our work.
Signature Path Prefetcher (SPP) is a lookahead prefetcher and consists
of following structural components:

\textbf{Signature Table:} The ST keeps track of 256 most recently accessed
pages.  It is meant to capture memory access patterns within a page
boundary.  SPP indexes into an entry of ST using the page number.  For
each entry corresponding to a page, ST stores the `last block offset'
and `old signature'.  Last block offset is the block offset of the
last memory access of that given page.  The block offset is calculated
with respect to the page boundary.  The signature is a 12-bit
compressed representation of the past few memory accesses for that
page.  The signature is calculated as:
$$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\; (\,Delta\,)$$ 
In case a matching page entry is found, the stored signature in the corresponding entry is retrieved.

%% NOTE: Can we do away with this example? 
%%       If it looks like we are weiting too much about SPP
Consider the case that incoming page number 10 with a block offset 3
finds a match in the ST.  The retrieved pattern signature is 0x30 and
the Last Offset is 1.  Since now there is the Last Block Offset (1),
Incoming Block Offset (3), Old Signature (0x30), and New Signature
calculated as per above equation (0x182), SPP can infer
non-speculatively that the given pattern of memory accesses (as
captured in the Old Signature) leads to the particular Delta.  In
general, Delta is defined as the difference between the prefetch
suggestion block and the initial block which triggered the prefetch.
In this case, since SPP is in learning phase, it is defined as the
difference between the Incoming Block Offset and the Last Block Offset
(+2 in this case).  This, in turn generates the new memory pattern
(New Signature).  This newly learned signature pattern and the delta
is stored in the Pattern Table.

\textbf{Pattern Table:} The PT is indexed by the signature generated
from ST.  PT holds predicted delta patterns and their confidence
estimates.  Each entry indexed by the signature holds up to 4 unique
delta predictions.  This is implemented by making PT as a 4-way
associative table.  To measure an estimate of a delta confidence, PT
implements two counters: C\textsubscript{sig} for each entry indexed
by the signature and C\textsubscript{delta} for each delta entry for a
given signature.  More details of this on-line score-keeping mechanism
is described under Confidence Tracking discussion.

\textit{Lookahead Prefetching:} On each trigger, SPP tries to go down
the program speculation path using its own prefetch suggestion.
Using current prefetch as the baseline, it re-accesses the PT to generate further
prefetches.  It keeps on repeating this cycle of accessing PT and
updating the signature based on highest confidence prefetch from the
last iteration.  These iteration count till which SPP manages to
predict prefetch entries in the lookahead manner is characterized as
its `depth'.  While doing this, SPP also keeps compounding the
confidence in each depth.  Thus as depth increases, overall confidence
keeps decreasing.

\textbf{Global History Register:} The GHR is a small 8-entry fully
associative structure to bootstrap learning across page boundaries. 
Whenever a prefetch suggestion from the
Pattern Table crosses page boundaries, instead of completely
discarding it, GHR saves that information for inter-page learning.
Since our NATCH implementation in this paper doesn't affect GHR
functionality in any ways, we won't be going into more details about
GHR.

\textbf{Prefetch Filter:} SPP also introduced the concept of Prefetch
Filter - a 1024 entry 1-way associative table which keeps a record of
last few entries prefetched.  This proves useful for following
reasons:
\begin{itemize}
\item \textbf{Eliminating duplicate prefetches}\newline Any subsequent
  prefetches recommended by SPP mechanism are checked against the
  existing entries in Prefetch Filter.  If a valid entry is found, it
  means that the particular cache line is already either in L2 Cache
  or being fetched.  Hence there is no need to issue a duplicate
  prefetch.

\item \textbf{Updating prefetcher}\newline Since the PF keeps a track
of past few prefetches, it uses that information to see if a given
prefetch led to a demand hit or a cache eviction. Using this information, 
SPP can update its internal states.

\item \textbf{Tracking Global Accuracy}\newline PF keeps a track of
  global usefulness ($\alpha$) of the prefetches by calculating 
  C\textsubscript{useful} / C\textsubscript{total}, where
  C\textsubscript{total} is the total number of prefetches
  getting recorded in the PF and C\textsubscript{useful} is the number 
  of prefetches that led to a demand hit.
\end{itemize}

\textbf{Confidence Tracking}: The PT keeps a track of hits to each
signature through a counter C\textsubscript{sig}.  The number of hits
for a given delta per signature are tracked using a counter
C\textsubscript{delta}.  The confidence for a given delta is
approximated through C\textsubscript{d} = C\textsubscript{delta} /
C\textsubscript{sig}.  When SPP enters into a lookahead mode, the path
confidence P\textsubscript{d} is given as:
$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;  P\textsubscript{d-1}$$ where $\alpha$ is called the global accuracy and is calculated as shown previously. 
The range of global accuracy ($\alpha$) is [0,1].

Here d is the lookahead depth.  For d = 1, when SPP is in
non-speculative mode, P\textsubscript{0} can be thought of as 1. 
The final P\textsubscript{d} is thresholded against prefetching-threshold
(T\textsubscript{p}) to decide whether to prefetch or not.  If yes,
then P\textsubscript{d} is thresholded against a numerically bigger
fill-threshold (T\textsubscript{f}) to decide whether the prefetch
should be sent to L2 Cache (high confidence prefetch) or Last Level
Cache (low confidence prefetch)

@@

Moreover, the on-line confidence mechanism introduced in SPP was very
rudimentary.  It was based on taking a ratio C\textsubscript{d} =
C\textsubscript{delta} / C\textsubscript{sig} as explained above.  The
same confidence was used to make the decision whether to prefetch or
not to prefetch; and which level to prefetch.  While this
approximation was shown to work in the original implementation, we believe that a
better form of generalised on-line decision making was possible.  
Hence, it was
necessary to build a robust and adaptable learning mechanism to accept
/ reject the prefetch suggestions; and to decide the fill level (L2
Cache vs Last Level Cache).

>>>>

%% NOTE: The below paper is literally the same idea as ours. But poor implementation and no good results. How much to talk about this paper?
Finally, the paper on Data Cache Prefetching with Perceptron
Learning~\cite{BadPerc} talks about the idea of two step prefetching.
The first step is an existing baseline prefetcher while the second
step is a perceptron based throttler.  The paper has lots of
limitations in terms of design and implementation.  That reflects in
the fact that it did not lead to any significant performance gain over
even the basic prefetchers like the Stride Prefetcher~\cite{Stride} and
the Markov Prefetcher~\cite{Markov}.

\subsection{Perceptron Learning in Architecture}
\label{Background-Perceptron}
Perceptron learning for computer architecture design has been around
for a while.  It was first popularized by Jimenez \textit{et. al.} for
doing branch prediction~\cite{Perc_Branch}. The biggest advantage of 
the perceptron model is that it can adjust its weight dynamically 
to increase / decrease the importance given to each feature - branch 
hostories in this case.
The structure that we use is derived from the model introduced in
Piecewise Linear Branch Prediction~\cite{Piece_Linear}.  Here the
feature itself is used to index into a hash of perceptron weights.
This form of indexing makes sure that the hyper-plane learned by the
perceptron weights is able to differentiate between linearly
inseparable outcomes.  \textit{<Is it true? Not sure>}

Perceptron Learning for Reuse Prediction~\cite{Perc_Reuse} extended the
hash perceptron architecture to perform prediction in context of cache
replacement.  Perceptron layer can learn to correlate the cache
replacement behaviour with a variety of features. The features
introduced in that paper were derived from Program Counters of the
current and the last few instructions and from the current memory 
access address.  This led to a
prediction mechanism that is highly accurate, has quick learning rate
and is adaptable to changing program behaviour.

@@

\subsection{Baseline Prefetcher: The Signature Path Prefetcher}
\label{Background-SPP}

While the basic idea of perceptron based prefetch filtering is applicable 
to any look-ahead prefetcher, we developed a practical implementation 
of the idea using SPP~\cite{SPP} as the underlying prefetcher. 
SPP is a simple, lightweight, lookahead prefetcher which can be easily 
modified to be highly aggressive.  
Here we describe the basic architecture of SPP.

\textbf{Signature Table:} The ST keeps track of 256 most recently accessed
pages.  It is meant to capture memory access patterns within a page
boundary.  SPP indexes into an entry of ST using the page number.  For
each entry corresponding to a page, ST stores the `last block offset'
and `old signature'.  Last block offset is the block offset of the
last memory access of that given page.  The block offset is calculated
with respect to the page boundary.  The signature is a 12-bit
compressed representation of the past few memory accesses for that
page.  The signature is calculated as:
$$New Signature = (\,Old Signature << 3 bits\,) \;\;XOR\;\; (\,Delta\,)$$ 
Delta is the numerical difference between the block offset of the 
current and the previous memory acces. In case a matching page entry 
is found, the stored signature retrieved and used to index into the 
Pattern Table.

\textbf{Pattern Table:} The PT is indexed by the signature generated
from ST.  PT holds predicted delta patterns and their confidence
estimates.  Each entry indexed by the signature holds up to 4 unique
delta predictions.  This is implemented by making PT as a 4-way
associative table.

%% NOTE: Can we do away with this example? 
%% If it looks like we are writing too much about SPP
%% EDIT: Decided to reorganize and remove this for now.
%% 	Consider the case that incoming page number 10 with a block offset 3
%% 	finds a match in the ST.  The retrieved pattern signature is 0x30 and
%% 	the Last Offset is 1.  Since now there is the Last Block Offset (1),
%% 	Incoming Block Offset (3), Old Signature (0x30), and New Signature
%% 	calculated as per above equation (0x182), SPP can infer
%% 	non-speculatively that the given pattern of memory accesses (as
%% 	captured in the Old Signature) leads to the particular Delta.  In
%% 	general, Delta is defined as the difference between the prefetch
%% 	suggestion block and the initial block which triggered the prefetch.
%% 	In this case, since SPP is in learning phase, it is defined as the
%% 	difference between the Incoming Block Offset and the Last Block Offset
%% 	(+2 in this case).  This, in turn generates the new memory pattern
%% 	(New Signature).  This newly learned signature pattern and the delta
%% 	is stored in the Pattern Table.

\textbf{Prefetch Filter:} The PF is a 1024 entry 1-way associative table 
which keeps a record of
last few entries prefetched.  This proves useful for updating the prefetcher 
states if a tracked prefetch leads to a demand hit or a cache eviction.

\textbf{Lookahead Prefetching:} On each trigger, SPP tries to go down
the program speculation path using its own prefetch suggestion.
Using current prefetch as the baseline, it re-accesses the PT to generate further
prefetches.  It keeps on repeating this cycle of accessing PT and
updating the signature based on highest confidence prefetch from the
last iteration.  These iteration count till which SPP manages to
predict prefetch entries in the lookahead manner is characterized as
its `depth'.  While doing this, SPP also keeps compounding the
confidence in each depth.  Thus as depth increases, overall confidence
keeps decreasing.

\textbf{Confidence Tracking}: The PT keeps a track of hits to each
signature through a counter C\textsubscript{sig}.  The number of hits
for a given delta per signature are tracked using a counter
C\textsubscript{delta}.  The confidence for a given delta is
approximated through C\textsubscript{d} = C\textsubscript{delta} /
C\textsubscript{sig}.  When SPP enters into a lookahead mode, the path
confidence P\textsubscript{d} is given as:
$$P\textsubscript{d} \;=\; \alpha  \;.\;  C\textsubscript{d}  \;.\;  P\textsubscript{d-1}$$ 
Here $\alpha$ represents the global accuracy, calculated as the ratio of 
the number of prefetches which led to a demand hit to the number of 
prefetches recommended in total. The range of $\alpha$ is [0,1].

`D' is the lookahead depth.  For d = 1, when SPP is in
non-speculative mode, P\textsubscript{0} can be thought of as 1. 
The final P\textsubscript{d} is thresholded against prefetching-threshold
(T\textsubscript{p}) to decide whether to prefetch or not.  If yes,
then P\textsubscript{d} is thresholded against a numerically bigger
fill-threshold (T\textsubscript{f}) to decide whether the prefetch
should be sent to L2 Cache (high confidence prefetch) or Last Level
Cache (low confidence prefetch)

@@

Moreover, the on-line confidence mechanism introduced in SPP was very
rudimentary.  It was based on taking a ratio C\textsubscript{d} =
C\textsubscript{delta} / C\textsubscript{sig} as explained previously. This 
confidence was used to make the decision whether to prefetch or
not to prefetch; and which level to prefetch.  While this
approximation was shown to work in the original implementation, we believe that a
better form of generalised on-line decision making was possible.  
Hence, it was
necessary to build a robust and adaptable learning mechanism to accept
/ reject the prefetch suggestions; and to decide the fill level (L2
Cache vs Last Level Cache).

### 2018-11-19 intro.tex
Updated Sections: Intro, Back, Method, Results

\section{INTRODUCTION}
\label{Introduction}
Since past few decades, processor and memory scaling has been enabled
with different philosophies in mind.  While processor scaling focused
on speed improvements, memory scaling was centred around increasing
the storage capacity.  The process technology reaching scaling
limitations led to Memory Wall~\cite{MemWall} - the increasing
performance gap between the processor speed and the memory speeds.
Various techniques have been developed to combat this widening gap.
Data prefetching is one such important technique.  Data prefetching
exploits the fact that in most applications, memory accesses have a
well-defined pattern.

An ideal prefetching scheme involves capturing this memory access
patterns to predict future prefetches in a timely manner.  Other
important decisions faced by the prefetcher are when to bring in the
prefetch block of data, which level of cache hierarchy to place the
block in and finally, what block to evict to accommodate the incoming
block.  Memory accesses patterns can be as simple as fetching the next
cache line or a complex but predictable pattern, like in pointer
chasing applications.  Predicting this often requires a trade-off
between coverage and accuracy.  For a prefetcher to capture highly
irregular access patterns, it needs to be highly aggressive (high
coverage).  That also means the prefetcher would end up recommending a
lot of useless prefetches, leading to cache pollution (low accuracy).
This effect would get even more pronounced in multi-core scenario
where the last level cache is a shared resource.  Conversely, a
conservative prefetcher would be highly accurate but might not make
enough useful predictions.

To this effect we propose NATCH: NeurAl prefeTCh tHrottler.  It
is an enhancement to the existing state of the art prefetcher and
overcomes the coverage vs accuracy trade-off in a robust manner.  The
idea is to use a state-of-the-art prefetcher as the base engine.
For that we choose Signature Path Prefetcher (SPP) and 
modify the design to make it as aggressive as it can get.
This enables the base engine to capture complex memory access patterns hence
leading to increased coverage.  This comes at a cost of increased DRAM
traffic, hence reduction in overall IPC.  The prefetch suggestions
recommended by SPP pass through the perceptron based throttler. Over
time, perceptron layer learns to correlate the prefetch recommendations 
with various available features, leading to elimination of
useless prefetches. This leads to an overall increased accuracy of
the prefetch scheme and reduction in DRAM traffic.

@@

\item Implementing the pereptron throttler in a robust and accurate
  practical prefetching mechanism.  The overall scheme yields
  increased accuracy as compared to the other state of the art
  prefetchers.  Moreover, the perceptron throttler learns to adapt
  itself to shared resource constraints - leading to an even increased
  performance in multi-core and lower capacity / bandwidth
  environments.

\item Examining the perceptron weights by doing a statistical study to
  quantify the contribution made by each feature.  We have also used
  this idea to develop a comprehensive set of features where each
  feature contributes some `new' information to the perceptron.  The
  details are explained under methodology, in Section \ref{Method-Overheads}.

\item Cross-validating the design across benchmarks.  All the
  development and parameter tuning was done on the SPEC 2017
  benchmark.  The results have been shown to be consistent across
  unseen benchmarks.  More details can be found in Section \ref{Results-CrossVal}.

@@

We tested NATCH on Champsim, an open source simulator which models
an Out-of-Order CPU using.  For evaluation, we used SPEC 2017
benchmarks.  On a single core configuration, NATCH increases
performance of the system by 39.5\% as compared to the baseline with
no prefetching, on the memory intensive subset of the benchmark.  This
outperforms the next best prefetcher, SPP, by 6.84\%.  On a multi-core
system running a mix of memory intensive subset of the SPEC 2017
traces, NATCH saw an improvement of 11.9\% over SPP for 4-core
system and 10.67\% for 8-core system.

In the results part in Section \ref{Results}, we show a detailed analysis of the
speedup achieved.  We also used SPEC 2006 and CloudSuite benchmarks
for cross-validation of the prefetcher implementation.

The rest of the paper is organized as follows.  Section \ref{Background} talks about
background work and recaptures the SPP architecture Section \ref{Design}
discusses the detailed implementation of the perceptron mechanism and
the features used.  Section \ref{Method} consists of the evaluation methodology
and explores the feature space for perceptron learning.  Section \ref{Results} is
the performance evaluation of our prefetcher and Section \ref{Conclusion} concludes
the paper.

>>>>

\section{INTRODUCTION}
\label{Introduction}
Processor scaling and memory scaling have been enabled
with different philosophies in mind.  While processor scaling focused
on speed improvements, memory scaling was centred around increasing
the storage capacity.  The process technology reaching scaling
limitations led to Memory Wall~\cite{MemWall} - the increasing
performance gap between the processor speed and the memory speeds.
Various techniques have been developed to combat this widening gap.
Data prefetching is one such important technique.  Data prefetching
exploits the fact that in most applications, memory accesses have a
well-defined pattern.

An ideal prefetching scheme involves capturing this memory access
patterns to predict future prefetches in a timely manner.  Other
important decisions faced by the prefetcher are when to bring in the
prefetch block of data, which level of cache hierarchy to place the
block in and finally, what block to evict to accommodate the incoming
block.  Memory accesses patterns can be as simple as fetching the next
cache line or a complex but predictable pattern, like in pointer
chasing applications.  Predicting this often requires a trade-off
between coverage and accuracy.  
Most prefetchers achieve this by maintaining an internal confidence 
mechanism by keeping a counters to track events like cache hits / miss, 
references to an entry in the table \textit{etc}. By thresholding this 
confidence against different values, the extent of aggressiveness 
can be adjusted.

For a prefetcher to capture highly
irregular access patterns, it needs to be highly aggressive (high
coverage).  That also means the prefetcher would end up recommending a
lot of useless prefetches, leading to cache pollution (low accuracy).
This effect would get even more pronounced in multi-core scenario
where the last level cache is a shared resource~\cite{Friendly}.  Conversely, a
conservative prefetcher would be highly accurate but might not make
enough useful predictions.

To this effect we propose NATCH: NeurAl prefeTCh tHrottler.  It
is an enhancement to the existing state of the art prefetcher and
overcomes the coverage vs accuracy trade-off in a robust manner.  The
idea is to use a state-of-the-art prefetcher as the base engine.
For that we choose Signature Path Prefetcher (SPP) and 
modify the design to make it as aggressive as it can get.
This enables the base engine to capture complex memory access patterns 
and go deeper into the speculation path. This
leading to increased coverage, which comes at a cost of increased DRAM
traffic.  The prefetch suggestions
recommended by the base prefetch engine pass through the perceptron based throttler. 
Over time, perceptron layer learns to correlate the prefetch recommendations 
with various available features, leading to elimination of
useless prefetches. This leads to an overall increased accuracy of
the prefetch scheme and reduction in DRAM traffic.

@@

\item Implementing the pereptron throttler in a robust and accurate
  practical prefetching mechanism, yielding a significant performance 
  improvement as compared to the other state of the art
  prefetchers.  Moreover, the perceptron throttler learns to adapt
  itself to shared resource constraints - leading to a further increased
  performance in multi-core and lower capacity environments.

\item Defining a methodology for determining an appropriate set of 
  features to be used regardless of the underlying prefetcher used.
  More details are explained in Section \ref{Method-Features}.

@@

On a single core configuration, NATCH increased
performance of the system by 6.84\% as compared to the 
next best prefetcher, SPP.  On a multi-core
system running a mix of memory intensive subset of the SPEC 2017
traces, NATCH saw an improvement of 11.9\% over SPP for 4-core
system and 10.67\% for 8-core system.

The rest of the paper is organized as follows.  Section \ref{Background} talks about
background work and recaptures the SPP architecture.  Section \ref{Design}
discusses the detailed implementation of the perceptron mechanism and
the features used.  Section \ref{Method} consists of the evaluation methodology
and explores the feature space for perceptron learning.  Section \ref{Results} is
the performance evaluation of our prefetcher and Section \ref{Conclusion} concludes
the paper.

### 2018-11-19 method.tex
Updated Sections: Intro, Back, Method, Results

\subsection{Performance Model}
\label{Method-Model}
For testing and comparing NATCH, we used ChampSim~\cite{Champsim}
simulator.  ChampSim is an enhanced version of the framework that was
used for the 2nd Data Prefetch Championship (DPC-2)~\cite{DPC_2}. We
model model a 1-core / 4-core / 8-core Out of Order machine. The
details of the configuration parameters is summarized in Table
\ref{tab:Sim_params}.

@@

The block size was fixed 64KB. Prefetching was only triggered at L2
Cache demand accesses but could be sent to L2 Cache or Last Level
Cache.  No L1 Data level prefetching was done.  LRU replacement policy
was used on all levels of cache hierarchies.  Branch prediction was
done using the perceptron branch predictor~\cite{Perc_Branch}.  Page
size was kept as 4KB.  Champsim operates all the prefetchers strictly
in the physical address space.

@@

We performed cross-validation of our NATCH model using SPEC
2006~\cite{SPEC2006} and CloudSuite~\cite{CloudSuite} benchmarks.  For
SPEC 2006, we developed 94 simpoints spread across the 29
applications.  For CloudSuite, we used the traces made available for
the 2nd Cache Replacement Competition (CRC-2)~\cite{CRC_2}.  The traces
include 4 multi-core applications with 6 distinct phases per
application.  In total, we used 285 traces representing workloads
across 53 applications.

\textit{Single-core performance:} For single-core simulations, we use
200 Million instructions to warm-up the microarchitectural structures
and next 1 Billion instructions to do detailed simulations and collect
run-time statistics.  We report the performance as speedup over
baseline \textit{i.e.}, no prefetching.  We take the Instruction Per
Cycle (IPC) of the prefetcher and divide it with the IPC of baseline.
The final number reported is the geometric mean of the speed-up
attained on individual traces.

\textit{Multi-core performance:} For multi-application workloads, we
generate 100 completely random mixes and another 100 mixes from the
memory intensive subset of SPEC 2017.  For 4-core workload, 200
Million instructions are used for warm-up and additional 1 Billion
instruction simulated for collecting statistics.  Each CPU keeps
executing its workload till the last CPU completes 1 Billion
instructions after warm-up.  For collecting IPC and other data, only
the first billion instructions are considered as the ``region of
interest''.

@@

We repeat the same process for 8-core workloads with the exception
that 20 Million warm-up instructions and 100 Million full instructions
were executed.  This was done so as to keep the simulation run-time
within reasonable limits as a single 8-core mix was taking up to 5
days to simulate 1 Billion instructions.

>>>>

\subsection{Performance Model}
\label{Method-Model}
For testing and comparing NATCH, we used ChampSim~\cite{Champsim}
simulator.  ChampSim is an enhanced version of the framework that was
used for the 2nd Data Prefetch Championship (DPC-2)~\cite{DPC_2}. We
model a 1-core / 4-core / 8-core Out of Order machine. The
details of the configuration parameters is summarized in Table
\ref{tab:Sim_params}.

@@

The block size was fixed 64KB. Prefetching was only triggered at L2
Cache demand accesses but could be directed to L2 Cache or Last Level
Cache. No L1 Data level prefetching was done. LRU replacement policy
was used on all levels of cache hierarchies. Branch prediction was
done using the perceptron branch predictor~\cite{Perc_Branch}.  Page
size was kept as 4KB.  Champsim operates all the prefetchers strictly
in the physical address space.

@@

\textit{Single-core performance:} For single-core simulations, we use
200 Million instructions to warm-up the microarchitectural structures
and next 1 Billion instructions to do detailed simulations and collect
run-time statistics. We report the IPC speed over
baseline \textit{i.e.}, no prefetching.
The final number reported is the geometric mean of the speed-up
attained on individual traces.

\textit{Multi-core performance:} For multi-application workloads, we
generate 100 random mixes and another 100 mixes from the
memory intensive subset of SPEC 2017.  For 4-core workload, 200
Million instructions are used for warm-up and additional 1 Billion
instruction simulated for collecting statistics.  Each CPU keeps
executing its workload till the last CPU completes 1 Billion
instructions after warm-up.  For collecting IPC and other data, only
the first billion instructions are considered as the ``region of
interest''.

@@

We repeat the same process for 8-core workloads, correspondingly with 16MB LLC.
Only exception is that 20 Million warm-up instructions and 100 Million 
full instructions were executed.  
This was done so as to keep the simulation run-time
within reasonable limits as a single 8-core mix was taking up to 3
days to simulate 1 Billion instructions.

\textit{Cross Validation:} We cross-validated our NATCH model using SPEC
2006~\cite{SPEC2006} and CloudSuite~\cite{CloudSuite} benchmarks.  
For single core
SPEC 2006, we developed 94 simpoints spread across all the 29
applications. For multi-core, 
we followed the same methodology as SPEC 2017. 
For CloudSuite, we used the traces made available for
the 2nd Cache Replacement Competition (CRC-2)~\cite{CRC_2}.  The traces
include 4 4-core applications with 6 distinct phases per
application.

In total, we used 285 traces representing workloads
across 53 applications. Throughout the paper for SPEC 2017, 
we consider memory intensive subset as the traces with LLC MPKI > \_\_. 
This includes 48 out of the 95 simpoints developed.
Since SPEC 2006 differs considerably in terms of memory behaviour, we 
define memory intensive subset as traces with LLC MPKI > \_\_.
This includes 47 out of 94 simpoints.

### 2018-11-19 results.tex
Updated Sections: Intro, Back, Method, Results

\subsection{Single-core Results}
\label{Results-Single}
Figure \_\_ shows the IPC speedup of the 4 prefetchers simulated.  All
the results have been normalized to baseline \textit{i.e.}, no
prefetching.  As can be seen, NATCH yields a speedup of XX\% over
the baseline.  This is equivalent to XX\% over DA-AMPM, xx\% over BOP
and XX\% over SPP.  Out of the XX benchmarks, NATCH nearly matches
or outperforms most of the prefetchers on XX traces.

At its peak, NATCH manages to get the IPC speedup of a factor of
over \textbf{3.4} on the trace \textit{602.gcc\_s-2226B}.  This also
corresponds to speedup gain of \textbf{22.4\%} over the next best
prefetcher - SPP.  In general, benchmarks \textit{602.gcc\_s,
  603.bwaves\_s, 605.mcf\_s, 621.wrf\_s} and \textit{649.fotonik3d\_s}
benefit the most from NATCH, as compared to SPP only prefetching.

@@

\textit{4-core environment} Figure \_\_ shows a comparison of speedups
obtained on workload mix of memory intensive subset of SPEC 2017.  We
plot all the 4 prefetchers, normalized to baseline.  The workloads
have been sorted in the increasing order of performance benefit, as
measured in terms of normalized weighted speedup.  As can be seen,
NATCH offers an IPC improvement by 33.5\% on these traces.  This is
5.5\% over the next best prefetcher SPP; and XX\% and YY\% over BOP
and DA-AMPM respectively.

This increased improvement by NATCH on a 4-core machine is not
surprising.  Since NATCH has an extremely intelligent throttler
working to eliminate useless prefetches before they can create cache
pollution.  This further amplifies the edge that SPP had over BOP.
With the advantage of an orthogonal prefetcher, SPP can be as
aggressive as possible and leave the accuracy aspect to perceptron
throttler.

\textit{8-core environment} To test the effectiveness of NATCH in a
tighter constrained multi-core environment, we simulated an 8-core
machine with 16 MB last Level Cache.  Other parameters were kept same
as discussed under Methodology in Section \ref{Method-Workloads}.  
\textit{<See if we can use
speed-up over BOP as a measure here>}

\subsection{Additional Memory Constraints}
\label{Results-AdditionalMem}
Figure \_\_ shows the performance of NATCH against the other
prefetchers in tighter constraint environments.  For sake of
simplicity, we show 2 traces where NATCH gives the most advantage
over baseline, 2 traces where NATCH doesn't perform as well as the
best prefetcher *<-replace with the prefetcher name once results are
available* and finally GeoMean across all the traces.

@@

The 4-core results for SPEC 2006 are \_\_.

>>>>

\subsection{Single-core Results}
\label{Results-Single}

Figure \_\_ shows the IPC speedup on the memory intensive subset of SPEC 2017. 
All the results have been normalized to baseline \textit{i.e.}, no
prefetching.  As can be seen, NATCH yields a speedup of XX\% over
the baseline.  This is equivalent to XX\% over DA-AMPM, xx\% over BOP
and XX\% over SPP.  Out of the XX benchmarks, NATCH nearly matches
or outperforms most of the prefetchers on XX traces.

At its peak, NATCH manages to get the IPC speedup of a factor of
over \textbf{XX} on the trace \textit{602.gcc\_s-2226B}.  This also
corresponds to speedup gain of \textbf{XX\%} over the next best
prefetcher - SPP.  In general, benchmarks \textit{602.gcc\_s,
  603.bwaves\_s, 605.mcf\_s, 621.wrf\_s} and \textit{649.fotonik3d\_s}
benefit the most from NATCH, as compared to SPP only prefetching.

On the full SPEC 2017 suite, NATCH improves the baseline by XX\%,
which is YY\% more than just SPP.

@@

\textit{4-core environment} Figure \_\_ shows a comparison of speedups
obtained on 4-core mix of memory intensive subset of SPEC 2017.  We
plot all the 4 prefetchers, normalized to baseline.  The workloads
have been sorted in the increasing order of IPC obtained with running NATCH.
As can be seen,
NATCH offers an IPC improvement by 51.7\% on these traces.  This is
11.9\% over the next best prefetcher SPP; and XX\% and YY\% over BOP
and DA-AMPM respectively.

On a different set of fully random SPEC 2017 4-core mixes, NATCH provides an IPC speedup of 
26.2\% over baseline, which is 5.7\% more than just SPP.

\textit{8-core environment} The sorted comparison of speedup obtained 
on the memory intensive 8-core mixes, is depicted in Fig \_\_.
NATCH improves baseline performance by 38.6\% and that is 10.7\% 
more than using just SPP.

For completely random SPEC 2017 mixes, NATCH improves performance by 
23.6\% over the baseline, corresponding to 4.8\% over SPP.

This increased improvement by NATCH on a multi-core machines is not
surprising.  Since NATCH has an extremely intelligent throttler, it ends up 
eliminating useless prefetches before they can create cache
pollution. This further amplifies the edge that SPP had over BOP in multi-core environment.

\textit{<See if we can use
speed-up over BOP as a measure here for showing scalability>}

\subsection{Additional Memory Constraints}
\label{Results-AdditionalMem}

\textit{Low-Bandwidth DRAM}
Figure \_\_(a) shows the performance of NATCH against other prefetchers. 
For sake of simplicity, we show 2 traces where NATCH gives the most advantage
over baseline, 2 traces where NATCH doesn't perform as well as the next
best prefetcher (replace with the prefetcher name once results are
available) and GeoMean across memory intensive and across all the 
SPEC 2017 traces. As can be seen, for the memory intensive subset,
 NATCH improves baseline by XX\% and that is XX\% more than using SPP.

\textit{Small LLC}
Figure \_\_(b) shows the similar analysis of NATCH.
For the memory intensive subset, a performance improvement of XX\% 
over baseline is seen with NATCH, XX\% more than just SPP.

@@

For 4-core memory intensive mixes, NATCH improves the baseline by 59.2\%, 
8.7\% ahead of SPP. For 8-core memory intensive mixes, the IPC speedup 
over baseline is 47.9\%, 11.5\% ahead of SPP.
